Wireless Automobile Detection, License Plate Processing, and Data Availability Network Proposal

### 2014-03-16 proposal/proposal.tex
Added summary

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators will monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists and maps as more lots are added to the system. The Association for Computing Machinery group (ACMx) at CSM has already started working on a system for monitoring lot capacities (see section 1.2). We will collaborate closely with the ACMx group, integrating our work with theirs to create and deploy a working system by the end of this semester (May 2014).

@@

The goal of this project is to work with the ACMx group to create and deploy the system in two parking lots on campus by the end of the semester. See section 2.7 for information about the deployment. The system itself will consist of several subsystems, which are described in detail in section 2.

@@

\subsection{License Plate Image Acquisition}
Once the Raspberry Pi receives the interrupt from the automobile detection system that a car is entering the lot it will begin the process of obtaining a picture of the cars license plate. This procedure will be separated into two separate sections, image acquisition and license plate detection.

\subsubsection{Image Acquisition}
While the automobile detection system correctly identifies when a vehicle enters the lot, the Raspberry Pi will be responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate in the picture. The timing difference between when the Raspberry Pi receives the interrupt and when it actually captures the picture will initially be determined experimentally using the average speed at which cars enter the parking lot. As the project progresses this timing will be fine tuned in order to capture the optimal image every time. The goal with this section is to only take one picture to capture the license plate, so as to minimize capture time and power consumption.

\subsubsection{License Plate Detection}
Once the image has been captured, the Raspberry Pi will then focus on detecting the license plate in the image and then cropping the image to only include the license plate. This will be done in order to minimize the size of the transmission and the power required for that transmission.

@@

\subsection{Central Base Station}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee Pro radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Because Raspbian is optimized for the Raspberry Pi, It should be easy to write some shell scripts that will take care of forwarding the data via HTTP to the server. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See section 2.6 for more detailed information about the error correction functionality of the web application.

@@

\subsubsection{Interface with Base Station Raspberry Pi}
The recognition and extraction of license plate characters will not directly interface with the Arduino Fio or Acquisition Raspberry Pi that takes the images of automobiles. Instead, it is assumed all communications will come through the base station Raspberry Pi, and that the base station Raspberry Pi will upload received images to the ACMx Linux server (See section 2.6.1 about the HTTP interface).

@@

\subsubsection{HTTP Interface}
The base station Raspberry Pi will need some way to communicate its data with the web application/MySQL database. The simplest way to meet this need is to create an HTTP interface on the server. We will need this interface to handle requests from the front end of the web application anyway, and so it makes sense to also allow the base station to communicate with our server through HTTP requests. We will create a REST-ful HTTP interface that corresponds closely to the database schema and the front-end of the webpage, and the base station will send all its data to the application via the same interface. For example, if an image of a license plate is detected, the base station would send an HTTP POST request to 'acmxlabs.org/parking/licenseImage' (actual urls may change when we start designing the interface), and the server would receive the request and send the image to the ALPR software, as described in section 2.5. When the web application needs to get the data associated with a license plate, it will send an HTTP GET request to 'acmxlabs.org/parking/licensePlate/123ABC'.

\subsubsection{Viewing License Plate Data}
As mentioned in section 2.5, the license plate images will be translated from an image into a string of plain text, and stored in the MySQL database on the ACMx server. One of the main functions of the web interface will be to display this data to administrators. The page that displays the license plate data will have a simple interface which will allow administrators to filter the data by date/time, lot, and license plate number. For example, an administrator should be able to type in a specific license plate number and see all the dates/times that plate has been seen, and in which lot. Alternatively, an administrator will be able to enter a date range and lot(s) and see all the license plates that entered/exited the lot(s) during that date range. For now we are only interested in a proof-of-concept type interface that shows we have the capability to display the collected plate data in a useful format; in the future it could be incredibly useful to cross reference our data with the campus parking department's database to check if a car has entered a lot without a permit.

\subsubsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, as mentioned in section 2.4, we will be collecting periodic overhead images from a camera mounted on the base station Raspberry Pi. These images will serve as a visual representation of how many cars are actually in the parking lot at the time the image was taken. Administrators should be able to periodically compare this image to the collected data from the sensor network to make sure the numbers being computed by the network accurately represent the number of cars in the lot. As previously mentioned, the deployment of this project will only be on two campus parking lots, and we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. The web interface will allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator should be able to overwrite the count in the database to reflect reality. This interface will be extremely useful in helping to calibrate the system. For example, if the sensors are consistently reporting less cars in the lot than the image shows, we can fine-tune the collection process to make sure fewer cars slip through unnoticed. Future work in this area will include an automated version of this process, where a computer program will compare the image to the sensed data and make any necessary corrections.

@@

\section{Summary}
A summary. Talk about the motivation behind this project. Why it is important.

>>>>

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators will monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists and maps as more lots are added to the system. The Association for Computing Machinery group (ACMx) at CSM has already started working on a system for monitoring lot capacities (see Section 1.2). We will collaborate closely with the ACMx group, integrating our work with theirs to create and deploy a working system by the end of this semester (May 2014).

@@

The goal of this project is to work with the ACMx group to create and deploy the system in two parking lots on campus by the end of the semester. See Section 2.7 for information about the deployment. The system itself will consist of several subsystems, which are described in detail in Section 2.

@@

\subsection{License Plate Image Acquisition}
Once the Acquisition Raspberry Pi receives the interrupt from the automobile detection system that an automobile is entering the lot it will begin the process of obtaining a picture of the vehicle's license plate. This procedure will be separated into two separate sections: image acquisition and license plate detection.

\subsubsection{Image Acquisition}
While the automobile detection system correctly identifies when a vehicle enters the lot, the Raspberry Pi will be responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate in the image. The timing difference between when the Raspberry Pi receives the interrupt and when it actually captures the picture will initially be determined experimentally using the average speed at which cars enter the parking lot. As the project progresses this timing will be fine tuned in order to capture the optimal image every time. The goal with this section is to only take one picture to capture the license plate, so as to minimize capture time and power consumption.

\subsubsection{License Plate Detection}
Once the image has been captured, the Raspberry Pi will then focus on detecting the license plate in the image and cropping the image to only include the license plate. This will be done in order to minimize the size of the transmission and the power required for that transmission.

@@

\subsection{Central Base Station}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee Pro radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing shell scripts that will take care of forwarding the data via HTTP to the server will be an easily achievable task. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See Section 2.6 for more detailed information about the error correction functionality of the web application.

@@

\subsubsection{Interface with Base Station Raspberry Pi}
The recognition and extraction of license plate characters will not directly interface with the Arduino Fio or Acquisition Raspberry Pi that takes the images of automobiles. Instead, it is assumed all communications will come through the base station Raspberry Pi, and that the base station Raspberry Pi will upload received images to the ACMx Linux server (See Section 2.6.1 about the HTTP interface).

@@

\subsubsection{HTTP Interface}
The base station Raspberry Pi will need some way to communicate its data with the web application/MySQL database. The simplest way to meet this need is to create an HTTP interface on the server. We will need this interface to handle requests from the front end of the web application anyway, and so it makes sense to also allow the base station to communicate with our server through HTTP requests. We will create a REST-ful HTTP interface that corresponds closely to the database schema and the front-end of the webpage, and the base station will send all its data to the application via the same interface. For example, if an image of a license plate is detected, the base station would send an HTTP POST request to 'acmxlabs.org/parking/licenseImage' (actual urls may change when we start designing the interface), and the server would receive the request and send the image to the ALPR software, as described in Section 2.5. When the web application needs to get the data associated with a license plate, it will send an HTTP GET request to 'acmxlabs.org/parking/licensePlate/123ABC'.

\subsubsection{Viewing License Plate Data}
As mentioned in Section 2.5, the license plate images will be translated from an image into a string of plain text, and stored in the MySQL database on the ACMx server. One of the main functions of the web interface will be to display this data to administrators. The page that displays the license plate data will have a simple interface which will allow administrators to filter the data by date/time, lot, and license plate number. For example, an administrator should be able to type in a specific license plate number and see all the dates/times that plate has been seen, and in which lot. Alternatively, an administrator will be able to enter a date range and lot(s) and see all the license plates that entered/exited the lot(s) during that date range. For now we are only interested in a proof-of-concept type interface that shows we have the capability to display the collected plate data in a useful format; in the future it could be incredibly useful to cross reference our data with the campus parking department's database to check if a car has entered a lot without a permit.

\subsubsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, as mentioned in Section 2.4, we will be collecting periodic overhead images from a camera mounted on the base station Raspberry Pi. These images will serve as a visual representation of how many cars are actually in the parking lot at the time the image was taken. Administrators should be able to periodically compare this image to the collected data from the sensor network to make sure the numbers being computed by the network accurately represent the number of cars in the lot. As previously mentioned, the deployment of this project will only be on two campus parking lots, and we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. The web interface will allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator should be able to overwrite the count in the database to reflect reality. This interface will be extremely useful in helping to calibrate the system. For example, if the sensors are consistently reporting less cars in the lot than the image shows, we can fine-tune the collection process to make sure fewer cars slip through unnoticed. Future work in this area will include an automated version of this process, where a computer program will compare the image to the sensed data and make any necessary corrections.

@@

\section{Summary}
This project extends the work done by ACMx to create a comprehensive parking management system for the Colorado School of Mines. The proposed expansions cover interfacing with the existing automobile detection network, acquiring images of license plates, processing those images, and presenting the resulting characters via a web application.

Acquiring images of license plates will include experimentally testing the timing of taking the picture, as well as on-mote pre-processing of the image. By completing some processing of the image on the Raspberry Pi, we can both reduce the size of data we need to transmit (saving battery power), while also avoiding transmitting false positives (when no license plate is detected).

Image acquisition will come from several Raspberry Pi's stationed at parking lot entrances. These images will be sent to a central base station, which will then upload the data to a server. Once on the server, additional image processing will be done to retrieve the characters composing the license plate. This data will be stored in a database for use in the web application.

In addition, the base station will also sample images of the parking lots in order to allow administrators to correct the status of the system. This may happen when the acquisition motes unsuccessfully identify a car entering or exiting, and therefore have incorrect information regarding the parking lot's capacity.

The web application will serve to show the state of the parking lots, as well as offer a way for administrators to reconcile observed data (from base station images) and sensed data (from acquisition motes). Administrators will also be able to access license plate data from processed images to show proof-of-concept for this utility.

Once deployed, the SmartLots program, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use this to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.
